 

### 2021-07-11 vertical-thermal-review.Rmd
Update vertical-thermal-review.Rmd

A few studies have found $Q_{10}$ to be greater in upper-canopy leaves than in shaded lower-canopy leaves [@turnbullScalingFoliarRespiration2003; @harleyEffectsLightTemperature1996], whereas @griffinCanopyPositionAffects2002 found that in Populas deltoides, lower canopy leaf respiration was more sensitive to changes in $T_{air}$ than upper canopy. Other studies have observed no difference in $Q_{10}$ within species vertical profiles and in seedling sun vs. shade leaves [@weerasingheCanopyPositionAffects2014; @bolstad_foliar_1999; @zaragoza-castellsDoesGrowthIrradiance2007; @zaragoza-castellsClimatedependentVariationsLeaf2008]. (*Martijn has some unpublished data on this that we might include.unpublished data showing higher Q10 in shade than in sun leaves for 10 species or so. May make sense to include here: https://github.com/EcoClimLab/vertical-thermal-review/issues/8.*). 
@xuSeasonalVariationTemperature2006 found that the canopy-position effect gradient in respiration rates was much greater for Quercus rubra trees in mesic than in xeric sites, but $Q_{10}$ did not differ between canopy positions.

>>>>

A few studies have found $Q_{10}$ to be greater in upper-canopy leaves than in shaded lower-canopy leaves [@turnbullScalingFoliarRespiration2003; @harleyEffectsLightTemperature1996], whereas @griffinCanopyPositionAffects2002 found that in *Populas deltoides*, lower canopy leaf respiration was more sensitive to changes in $T_{air}$ than upper canopy. Other studies have observed no difference in $Q_{10}$ within species vertical profiles and in seedling sun vs. shade leaves [@weerasingheCanopyPositionAffects2014; @bolstad_foliar_1999; @zaragoza-castellsDoesGrowthIrradiance2007; @zaragoza-castellsClimatedependentVariationsLeaf2008]. (*Martijn has some unpublished data on this that we might include.unpublished data showing higher Q10 in shade than in sun leaves for 10 species or so. May make sense to include here: https://github.com/EcoClimLab/vertical-thermal-review/issues/8.*). 
@xuSeasonalVariationTemperature2006 found that the canopy-position effect gradient in respiration rates was much greater for Quercus rubra trees in mesic than in xeric sites, but $Q_{10}$ did not differ between canopy positions.
